Automatic prostate cancer detection through \acs*{dce}-\acs*{mri} images: all you need is a good normalization

### 2016-08-24 03_experiments/experiments.tex
Finish the first draft

\begin{table}
  \caption{\acs*{auc} for each individual pharmacokinetic parameter using a \acs*{nb} classifier.}
  \centering
  \begin{tabular}{l c c}
    \toprule
    \textbf{Features} & Un-normalized data & Normalized data \\
    \midrule
    \textbf{Brix model} & & \\
    \quad $A$         & 0.62 & 0.67 \\
    \quad $k_{el}$    & 0.52 & 0.61 \\
    \quad $k_{ep}$    & 0.52 & 0.58 \\
    \textbf{Hoffmann model} & & \\
    \quad $A$         & 0.50 & 0.56 \\
    \quad $k_{el}$    & 0.53 & 0.64 \\
    \quad $k_{ep}$    & 0.50 & 0.66 \\
    \textbf{Tofts model with population \ac{aif}} & & \\
    \quad $K_{trans}$ & 0.62 & 0.65 \\
    \quad $v_{e}$     & 0.50 & 0.52 \\
    \quad $v_{p}$     & 0.63 & 0.53 \\
    \textbf{Tofts model with patient \ac{aif}} & & \\
    \quad $K_{trans}$ & 0.66 & 0.65 \\
    \quad $v_{e}$     & 0.50 & 0.52 \\
    \quad $v_{p}$     & 0.37 & 0.65 \\
    \textbf{\ac{pun} model} & & \\
    \quad $a_0$       & 0.53 & 0.51 \\
    \quad $r$         & 0.59 & 0.55 \\
    \quad $\beta$     & 0.56 & 0.44 \\
    \textbf{Semi-quantitative analysis} & & \\
    \quad wash-in     & 0.64 & 0.51 \\
    \quad wash-out    & 0.50 & 0.66 \\
    \quad IAUC        & 0.61 & 0.64 \\
    \quad $\tau$      & 0.57 & 0.61 \\
    \quad $S_M - S_0$ & 0.63 & 0.64 \\
    \bottomrule
  \end{tabular}
  \label{tab:resfeats}
\end{table}

This experiment consists in assessing the benefit of normalizing the \ac{dce}-\ac{mri}, prior to classify each individual parameter for each model.
Therefore, a Gaussian \ac{nb} classifier is used in conjunction with a \ac{lopo}.
The results are summarized in Table~\ref{tab:resfeats}.
In general, data normalization leads to better classification performance while extracting pharmacokinetic related parameters.
Only the \ac{pun} model does not follow this tendency.
{\color{red} Find out why!!!}

@@

\begin{figure}
  \centering
  \hspace*{\fill}
  \subfigure[Without normalization.]{\label{fig:rfpharmaunorm}\includegraphics[width=.49\textwidth]{03_experiments/figures/unormalized/rf.pdf}} \hfill
  \subfigure[With normalization.]{\label{fig:rfpharmanorm}\includegraphics[width=.49\textwidth]{03_experiments/figures/normalized/rf.pdf}}
  \hspace*{\fill}
  \caption{\acs*{roc} analysis using a \acs*{rf} classifier with and without normalization \ac{dce}-\ac{mri} data for different pharmacokinetic models.}
  \label{fig:normpharmarf}
\end{figure}

In previous proposed \ac{cad} systems, the pharmacokinetic of a single model are combined.
Therefore, for each model, all parameters are combined and classified using a \ac{rf} classifier in a \ac{lopo} fashion.
The use of \ac{rf} is motivated since that it leads to the best performance in the state-of-the-art methods~\citep{litjens2014computer}.
Results are summarized by performing a \ac{roc} analysis and computing the \ac{auc}, as reported in Fig.\,\ref{fig:normpharmarf}.
Independently to the quantification model, the data normalization improves the\ac{auc} and classification performance.

@@

As stated in the introduction, the quantification methods are extracting a set of parameters characterizing the enhancement \ac{dce}-\ac{mri} signal.
However, this extraction could lead to a loss of information.
This experiment is performed to assess if making use of the whole \ac{dce}-\ac{mri} signal instead of the just the pharmacokinetic parameters can improve the classification performance.
Therefore, each enhanced \ac{dce}-\ac{mri} signal, normalized and un-normalized, is classified using a \ac{rf} classifier in a \ac{lopo} fashion.
The \ac{roc} analysis and \ac{auc} are reported in Fig.\,\ref{fig:rfnormdcesignal}.
It can be noted that the worst performance are achieved while classifying raw data without normalization.
However, data normalization in conjunction with the use of the whole \ac{dce}-\ac{mri} signal is the strategy which outperforms all others.

>>>>

\begin{table*}
  \caption{\acs*{auc} for each individual pharmacokinetic parameter using a \acs*{nb} classifier.}
  \centering
  \begin{tabular}{l c c c}
    \toprule
    \textbf{Features} & Un-normalized data & Normalized data & Gain/Loss \\
    \midrule
    \textbf{Brix model} & & & \\
    \quad $A$         & 0.62 & 0.67 & + 0.05 \\
    \quad $k_{el}$    & 0.52 & 0.61 & + 0.09 \\
    \quad $k_{ep}$    & 0.52 & 0.58 & + 0.06 \\
    \textbf{Hoffmann model} & & & \\
    \quad $A$         & 0.50 & 0.56 & + 0.06 \\
    \quad $k_{el}$    & 0.53 & 0.64 & + 0.11 \\
    \quad $k_{ep}$    & 0.50 & 0.66 & + 0.16 \\
    \textbf{Tofts model with population \ac{aif}} & & & \\
    \quad $K_{trans}$ & 0.62 & 0.65 & + 0.03 \\
    \quad $v_{e}$     & 0.50 & 0.52 & + 0.02 \\
    \quad $v_{p}$     & 0.63 & 0.53 & - 0.10 \\
    \textbf{Tofts model with patient \ac{aif}} & & & \\
    \quad $K_{trans}$ & 0.66 & 0.65 & - 0.01 \\
    \quad $v_{e}$     & 0.50 & 0.52 & + 0.02 \\
    \quad $v_{p}$     & 0.37 & 0.65 & + 0.18 \\
    \textbf{\ac{pun} model} & & & \\
    \quad $a_0$       & 0.53 & 0.51 & - 0.02 \\
    \quad $r$         & 0.59 & 0.55 & - 0.04 \\
    \quad $\beta$     & 0.56 & 0.44 & - 0.12 \\
    \textbf{Semi-quantitative analysis} & & & \\
    \quad wash-in     & 0.64 & 0.51 & - 0.15 \\
    \quad wash-out    & 0.50 & 0.66 & + 0.16 \\
    \quad IAUC        & 0.61 & 0.64 & + 0.03 \\
    \quad $\tau$      & 0.57 & 0.61 & + 0.04 \\
    \quad $S_M - S_0$ & 0.63 & 0.64 & + 0.01 \\
    \bottomrule
  \end{tabular}
  \label{tab:resfeats}
\end{table*}

This experiment consists in assessing the benefit of normalizing the \ac{dce}-\ac{mri}, prior to classify each individual parameter for each model.
Therefore, a Gaussian \ac{nb} classifier is used in conjunction with a \ac{lopo}.
The results are summarized in Table~\ref{tab:resfeats}.
In general, data normalization leads to better classification performance (i.e., + 4\%) while extracting pharmacokinetic related parameters.
Only the \ac{pun} model does not follow this tendency.
{\color{red} Find out why!!! I am computing the R2 coefficient to check the fit of data.}

@@

\begin{figure*}
  \centering
  \hspace*{\fill}
  \subfigure[Without normalization.]{\label{fig:rfpharmaunorm}\includegraphics[width=.49\textwidth]{03_experiments/figures/unormalized/rf.pdf}} \hfill
  \subfigure[With normalization.]{\label{fig:rfpharmanorm}\includegraphics[width=.49\textwidth]{03_experiments/figures/normalized/rf.pdf}}
  \hspace*{\fill}
  \caption{\acs*{roc} analysis using a \acs*{rf} classifier with and without normalization \ac{dce}-\ac{mri} data for different pharmacokinetic models.}
  \label{fig:normpharmarf}
\end{figure*}

In previous proposed \ac{cad} systems, the pharmacokinetic of a single model are combined.
Therefore, for each model, all parameters are combined and classified using a \ac{rf} classifier in a \ac{lopo} fashion.
The use of \ac{rf} is motivated since that it leads to the best performance in the state-of-the-art methods~\citep{litjens2014computer}.
Results are summarized by performing a \ac{roc} analysis and computing the \ac{auc}, as reported in Fig.\,\ref{fig:normpharmarf}.
Independently to the quantification model, the data normalization improves the\ac{auc} and classification performance.
An increase of the \ac{auc} ranging from 1\% to 12\% is noticeable.

@@

As stated in the introduction, the quantification methods are extracting a set of parameters characterizing the enhancement \ac{dce}-\ac{mri} signal.
However, this extraction could lead to a loss of information.
This experiment is performed to assess if making use of the whole \ac{dce}-\ac{mri} signal instead of the just the pharmacokinetic parameters can improve the classification performance.
Therefore, each enhanced \ac{dce}-\ac{mri} signal, normalized and un-normalized, is classified using a \ac{rf} classifier in a \ac{lopo} fashion.
The \ac{roc} analysis and \ac{auc} are reported in Fig.\,\ref{fig:rfnormdcesignal}.
Classification without normalization lead to the worst performance, with an \ac{auc} of 0.57.
However, data normalization in conjunction with the use of the whole \ac{dce}-\ac{mri} signal is the strategy which outperforms all others, with an \ac{auc} of 0.67.

### 2016-08-24 04_discussions/discussions.tex
Finish the first draft

\section{Experiments and results}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../main"
%%% End:

>>>>

\section{Discussions}\label{sec:discussions}

The experiments conducted in the previous section can give rise to several discussions.
In Tofts quantification, two different approaches have been used to infer the pharmacokinetic parameters: using a population-based or a patient-based \ac{aif}.
The population-based \ac{aif} approach leads to better classification performance.
However, there is two shortcomings to take into account while advancing this fact:
(i) T$_10$ parameter has been fixed and not computed from a T$_1$ map and
(ii) the population-based \ac{aif} has been estimated from a cohort of only 17 patients.
These two limitations have to be considered while advancing that population-based \ac{aif} modelling is outperforming patient-based \ac{aif} modelling.

The best classification performance is reached by normalizing the \ac{dce}-\ac{mri} data and use the whole enhanced signal as feature, emphasizing the fact that a loss of information while extracting quantitative parameters.
However, this strategy suffers from one drawback: the training time of the \ac{rf} classifier increases since that from 3 to 5 features, the feature space become a 40 dimensions space.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "../main"
%%% End:

### 2016-08-24 05_conclusions/conclusions.tex
Finish the first draft

\section{Discussions}

>>>>

\section{Conclusions and future works}\label{sec:conclusions}

In this work, we presented a new method for normalizing/standardizing \ac{dce}-\ac{mri} data.
This method aimed at reducing the inter-patient variations occurring during data acquisition.
A graph-based approach was used to correct intensity offset in conjunction with a model-based correction to reduce time offset as well as intensity scaling.
We show the benefit of our normalization method prior to extract quantitative and semi-quantitative features, with a significant improvement of the classification performance.
Nevertheless, we also show that using the whole normalized \ac{dce}-\ac{mri} signal outperforms all quantitative approach.

As avenues for future research, this normalization has to be part of a \ac{mpmri} \ac{cad} system in which \ac{dce}-\ac{mri} modality needs to be combined with other complementary modalities.
